MGRFE: multilayer recursive feature elimination based on an embedded genetic algorithm for cancer classification

### 2018-12-01 MGRFE Response 2018-12-01/Latex Manuscript/response.txt

Response: Thank you for this comment. Our proposed approach has showed great efficiency in gene selection through experimental comparisons on large amount of high-dimensional expression datasets with the most state-of-the-art algorithms. The chief innovation of MGRFE is combining the evolutionary strategy of GA with recursive feature elimination method, and the results of MGRFE is better than methods based only on recursive feature elimination method or only genetic algorithm respectively.
The RFE feature selection method has advantage of high execution speed, meanwhile limitation of dissatisfactory classification accuracy for gene selection. The swarm intelligence based gene selection approaches have advantage of powerful heuristic search ability in finding optimal gene subset, but shortcomings such as slow convergence speed and existing of irrelevant features in selected feature subset. Through combining the evolution calculation of genetic algorithm and the explicit feature elimination of RFE process, the designed MGRFE can take the best points of these two kind of methods and avoid their limitations, thus could find the minimal discriminatory gene subset with high convergence speed.
According to the though experiments performed on the 19 benchmark datasets, the proposed MGRFE could offer smaller informative gene subsets but the same or higher phenotype diagnosis accuracies compared with the current state-of-the-art feature selection methods.

3.	The authors claim to compare their method on 17 data sets. But I did not see any evidence that the finally determined feature set is validated on an independent data set of the same form of cancer for example. All that the authors have done is five-fold cross-validation within the same data set.  Without this sort of validation on an independent data set, the claimed performance figures by themselves are not very persuasive. This is because cross-validation within the same data set does not take into account factors such as batch effect, platform variation, and the like.

Response: Thank you for this comment.
1.	There are two main difficulties for validate the determined feature set on independent gene expression data set.
a)	The very limited available benchmark datasets for one typical disease. It is difficult to acquire sufficient and appropriate bio-samples due to high expense of micro-array sample collection and other various factors[13], thus the available benchmark datasets are limited and the sample number in each data set is usually small. For many diseases, we just have one widely used microarray benchmark, like the colon cancer (Colon)[14] and small round blue cell tumors (SRBCT)[15].
b)	For microarray benchmark datasets about same disease, the features and sample classes are usually different. Different microarray datasets usually have different gene features for the gene probes vary among different microarray analysis platform. For example, on the leukemia related datasets of Leuk and MLL used in this study, the gene probes are very different for generating from different microarray platforms.
Thus, the currently published gene selection algorithms on microarrays are commonly validated within each microarray benchmark dataset.
2.	We manage to validate the selected gene subsets of Leuk and Gas1/Gas2 on independent datasets as shown in Table 4. The results of independent feature subset validation are added to the “S8” section in Supplementary Material. The Table 4 in this response document is the Table 7 in the supplementary section “S8” .
a)	The leukemia datasets Leuk and MLL both have the sample data for ALL (acute lymphoblastic leukemia) and AML(acute myeloid). First, on Leuk the selected gene probes are [M23197, M31523]. Second, the genes related to these two probes are [CD33, TCF3]. Third, for these two genes, the gene probes in MLL are [32874_at, 36802_at, 1373_at, 1374_g_at]. Finally, we test the classification performance of the obtained 4 gene probes for ALL and AML samples in MLL dataset.
b)	Gas1 is about non-cardia gastric cancer, and Gas2 is about cardia gastric cancer. These two datasets are both from ref. [16] and have the same gene probes as features. The gene probes selected by MGRFE on Gas1 are [215380_s_at, 221928_at, 214746_s_at], and the gene probes selected on Gas2 are [210125_s_at, 206361_at]. We validated the gene probe subsets on the other dataset.

Table 4. Independent validation of features selected on Leuk and Gas1/Gas2 by 10-time 10-fold cross validation.
Feature test on	Feature from	Sn	Sp	Acc	Avc	MCC	AUC
MLL	Leuk	0.963	0.96	0.963	0.962	0.934	0.993
MLL	MLL	1	1	1	1	1	1
Leuk	Leuk	0.99	1	0.993	0.995	0.987	1

Gas1	Gas1	0.984	0.965	0.974	0.974	0.952	0.989
Gas1	Gas2	0.917	0.929	0.923	0.923	0.853	0.967
Gas2	Gas1	0.933	0.827	0.880	0.880	0.774	0.973
Gas2	Gas2	1	1	1	1	1	1
For dataset MLL, only ALL and AML samples are taken into consideration in this experiment to stay consistent with Leuk dataset.

From Table 3, it can be noted that the selected gene features on Leuk achieved satisfying performance on MLL. The obtained accuracy is 0.963, just slightly lower than the classification accuracies achieved within the datasets MLL or Leuk. The gene subset in Gas1 and Gas2 also showed acceptable performance on the other dataset. The different gastric cancer subtypes could account for the performance decrease in these two datasets.

4.	The authors' preferred method of genetic algorithms is known to lack theoretical foundations, to be very sensitive to various parameters in the algorithm, and to be extremely time consuming.  In contrast, the original paper where RFE was proposed, by Isabel Guyon, used the support vector machine (SVM) which is very fast and for which lots of theoretical results are available.  This is another reason for my not being overly enthusiastic about the paper.

Response: Thank you for this comment.
Firstly, for selecting informative gene features in a microarray, the state-of-the art methods are commonly evolutionary-computation based. Although the SVM-RFE method has many theoretical results, the classification accuracy of generated gene subset is likely to be lower than the result of evolutionary-computation based methods. The currently published leading methods of gene selection in microarray are usually base on swarm intelligence algorithms[17-19].
Secondly, there are several limitations of the RFE method which could not be ignored: a). the weights ranking could not exactly and completely reflect the importance of each gene; b). the top-ranked genes do not mean the best gene subset. Based on our experiment results, genes should be selected in combination but not individually; and c). there is no opportunity for a gene to appear again after being removed. On the contrast, the proposed MGRFE has been well-designed to avoid the above limitations by introducing the evolution computation strategy, thus has more advantages in finding the minimal informative gene subset. Fu and Fu-Liu evaluated SVM-RFE on datasets SRBCT and ALL AML and finally selected 19 and 4 genes to achieve 100% and 97.6% test accuracies, respectively [20]. But MGRFE selected only 5 and 2 genes to attain 100% accuracies in 5-fold CV for the same datasets.
Thirdly, compared with existed GA algorithm, the introduced RFE process has significantly enhanced the convergence speed and reduced running time. Instead of relying on widely used binary encoding, our proposed method utilizes variable length integer encoding in GA and cuts down the encoding length recursively in search process, which could quickly remove the irrelevant and redundant features and converge to the minimal informative feature combination. Kar et al. employed the evolutionary computation method PSO to select gene subset on three datasets SRBCT, ALL AML, and MLL. Their PSO-based method cost 2.7956, 2.7906 and 7.1488 hours on the three datasets respectively. In contrast, MGRFE merely used 10.8230, 9.0108 and 8.8739 minutes respectively in the same three datasets. Moreover, the selected gene subsets by MGRFE are smaller but with same or higher classification accuracies compared with Kar et al.’s PSO based method.
Fourthly, time complexity is of secondary significance in this issue, what should be prioritized is the discriminating ability of selected gene subset. For each microarray data set, just one running of the feature selection method is enough to generated the informative genes and minimal gene feature subset, which would be used repeatedly in the later classification or clustering applications. Thus, the running time of feature selection method is less important than its ability to locate the discriminatory genes.

5.	There are several places where the authors do not appear to be aware of simple statistical facts.  For instance, the accuracy is a weighted average of the sensitivity and the specificity.  But the authors talk as though they are independent parameters.  Equation (1) in the right column of page 1 is too wide.

Response: Thank you for this comment. In the revised manuscript, we have made effort to modify and optimize the expressions about statistical terms. Also, the format of our revised manuscript has been adjusted and we make the equations more organized.

6.	In Section 2.3.1 the authors use the T-test and MIC to achieve a first-cut reduction in the feature set.  I have found that using the so-called "volcano plot" which combines the T-test with a fold-change criterion, works better than just the T-test alone.

Response: Thank you for this comment. Since the “volcano plot” can combine the advantages of t-test and fold-change (FC), we use the “volcano plot” to replace the t-test and performed comparison experience on 2 balanced datasets (Adeno and Pros) and 2 imbalanced datasets (DLBCL and Leuk) by 5-fold cross validation. The results are shown in Table 5. According to experiment results, these two methods select same size of genes and achieve similar performance on all the tested 4 datasets. 
In our experiments, we noted one defect of “volcano plot” for gene selection in a number of different microarray datasets. When we use the “volcano plot” to selected informative genes, in each microarray dataset, we need to hand-tune the p-value threshold in t-test and especially fold-change threshold value in FC to obtain the satisfactory result. For example, on dataset Adeno, there are 894 informative genes have FC value larger or equal to 2; but on dataset Pros, the highest FC value in all genes is just 1.46. Thus, for different datasets, the threshold values in “volcano plot” should be different. In fact, for each of the tested 4 datasets, the threshold values in “volcano plot” have been hand-tuned individually and the finally assigned threshold values vary among different datasets. This situation pose difficulty for building automatically application on the wide range of microarray datasets. In contrast, the t-test has the consistent p-value setting in all the 17 datasets in experiments and shows more convenience. Due to the length restriction of the main manuscript, we added the filter method combination results to the “S7” section in Supplementary Material. The Table 5 becomes part of the Table 6 in the supplementary section “S7”.

Table 5. Compare the gene selection method of t-test and "volcano plot" on 2 balanced datasets (Adeno and Pros) and 2 imbalanced datasets (DLBCL and Leuk) using 5-fold cross validation.

Filter Methods	Dataset	Genes	Sn	Sp	Acc	Avc	MCC	AUC
Volcano plot
+MIC	Adeno	1	1	1	1	1	1	1
	Pros	4	0.980	0.982	0.980	0.981	0.963	0.968
	DLBCL	3	1	1	1	1	1	1
	Leuk	2	1	1	1	1	1	1
t-test
+MIC	Adeno	1	1	1	1	1	1	1
	Pros	4	0.980	0.982	0.981	0.981	0.963	0.980
	DLBCL	3	1	1	1	1	1	1
	Leuk	2	1	1	1	1	1	1

Reference

1.	Auger, N., C. Nicaud, and C. Pivoteau, Merge Strategies: from Merge Sort to TimSort. 2015.
2.	Fisher, R.A., Statistical Methods for Research Workers. 1958: Oliver and Boyd. 66-70.
3.	Cui, X.Q. and G.A. Churchill, Statistical tests for differential expression in cDNA microarray experiments. Genome Biology, 2003. 4(4).
4.	Lazar, C., et al., A Survey on Filter Techniques for Feature Selection in Gene Expression Microarray Analysis. Ieee-Acm Transactions on Computational Biology and Bioinformatics, 2012. 9(4): p. 1106-1119.
5.	Shen, Q., W.M. Shi, and W. Kong, Hybrid particle swarm optimization and tabu search approach for selecting genes for tumor classification using gene expression data. Computational Biology and Chemistry, 2008. 32(1): p. 53-60.
6.	Ge, R.Q., et al., McTwo: a two-step feature selection algorithm based on maximal information coefficient. Bmc Bioinformatics, 2016. 17.
7.	Reshef, D.N., et al., Detecting Novel Associations in Large Data Sets. Science, 2011. 334(6062): p. 1518.
8.	Mohamad, M.S., An enhancement of binary particle swarm optimization for gene selection in classifying cancer classes. Algorithms for Molecular Biology, 2013. 8(1): p. 1-11.
9.	Chuang, L., et al., Improved binary PSO for feature selection using gene expression data. Computational Biology & Chemistry, 2008. 32(1): p. 29-38.
10.	Li-Yeh, C., Y. Cheng-Huei, and Y. Cheng-Hong, Tabu search and binary particle swarm optimization for feature selection using microarray data. Journal of Computational Biology A Journal of Computational Molecular Cell Biology, 2009. 16(12): p. 1689.
11.	Li-Yeh, C., et al., A hybrid BPSO-CGA approach for gene selection and classification of microarray data. Journal of Computational Biology A Journal of Computational Molecular Cell Biology, 2012. 19(1): p. 68.
12.	Moosa, J.M., et al., Gene selection for cancer classification with the help of bees. Bmc Medical Genomics, 2016. 9(Suppl 2): p. 47.
13.	Dougherty, E.R., Small sample issues for microarray-based classification. Comparative and Functional Genomics, 2001. 2(1): p. 28-34.
14.	Alon, U., et al., Broad patterns of gene expression revealed by clustering analysis of tumor and normal colon tissues probed by oligonucleotide arrays. Proceedings of the National Academy of Sciences of the United States of America, 1999. 96(12): p. 6745-6750.
15.	Khan, J., et al., Classification and diagnostic prediction of cancers using gene expression profiling and artificial neural networks. Nature Medicine, 2001. 7(6): p. 673-679.
16.	Wang, G.S., et al., Comparison of Global Gene Expression of Gastric Cardia and Noncardia Cancers from a High-Risk Population in China. Plos One, 2013. 8(5).
17.	Han, F., et al., A Gene Selection Method for Microarray Data Based on Binary PSO Encoding Gene-to-Class Sensitivity Information. Ieee-Acm Transactions on Computational Biology and Bioinformatics, 2017. 14(1): p. 85-96.
18.	Motieghader, H., et al., A hybrid gene selection algorithm for microarray cancer classification using genetic algorithm and learning automata. Informatics in Medicine Unlocked, 2017. 9: p. 246-254.
19.	Dashtban, M. and M. Balafar, Gene selection for microarray cancer classification using a new evolutionary method employing artificial intelligence concepts. Genomics, 2017. 109(2): p. 91-107.
20.	Fu, L.M. and C.S. Fu-Liu, Evaluation of gene importance in microarray data based upon probability of selection. Bmc Bioinformatics, 2005. 6.
